### 2014-04-02 reanalysis/Fadi_Towfic_generic_medicine/towfic_et_al.rmd
figure script added

# the mean uses log2 numbers and will be a geometric mean.
meanM = rowMeans(thisdata[, sampleannotation[, "covariate"]=="M"]) 
meanRS = rowMeans(thisdata[, sampleannotation[, "covariate"]=="RS"]) 
meanDP = rowMeans(thisdata[, sampleannotation[, "covariate"]=="DP"])
meanN = rowMeans(thisdata[, sampleannotation[, "covariate"] =="N"])

foldchange_RS_vs_M = meanRS - meanM
foldchange_RS_vs_M = foldchange_RS_vs_M[meanM>6 | meanRS>6]
top1000_RS_vs_M = names(foldchange_RS_vs_M[order(abs(foldchange_RS_vs_M), decreasing=TRUE)])[1:1000]
table(table_s2$ID %in% top1000_RS_vs_M) #  851 out of 909. For some reason it is not a 100% match.
```
Out of Table S2's **`r nrow(table_s2)`** probes, **`r sum(table_s2$ID %in% top1000_RS_vs_M)`** were reproduced. Scatter plots of the actual values shows a correlation.

@@

```{r dev='png', fig.width=8, fig.height=8, fig.show='hold', tidy=FALSE}
group = factor(sampleannotation$covariate)
block = as.factor(sampleannotation$chip)
design = model.matrix(~0+group+block)
fit = lmFit(thisdata, design)
cont.matrix = makeContrasts ( contrasts="groupDP-groupN", levels=design)  
fit2 = contrasts.fit(fit, cont.matrix)
limma_ret_woc = eBayes(fit2)
limma_p_woc = limma_ret_woc$p.value
names(limma_p_woc) = names(limma_ret_woc$Amean)
rm(design, group, block, fit, cont.matrix, fit2)
table(p.adjust(limma_p_alt , method="fdr")<0.05, p.adjust(limma_p_woc , method="fdr")<0.05, dnn=c("ComBat adj", "LIMMA adj"))
```
The number of significant probes using the given FDR threshold of 0.05 when batch effect is handled by LIMMA is 
**`r sum(p.adjust(limma_p_woc , method="fdr")<0.05)`**. This is much lower compared to when ComBat is applied, 
**`r sum(p.adjust(limma_p_alt , method="fdr")<0.05)`**.  The p-value distribution is still skewed, but not as much.
```{r dev='png', fig.width=8, fig.height=8, fig.show='hold', tidy=FALSE}
hist(limma_p_alt, breaks=100, freq=T, border=thiscolors[1], main="Reproduced ANOVA p-values for DP vs. N")
hist(limma_p_woc, breaks=100, add=T, freq=T, border=thiscolors[2])
legend("topright", legend=c("ComBat adjusted", "LIMMA adjusted"), text.col=thiscolors)
```

>>>>

# the mean uses log2 numbers and will be a geometric mean.
meanM = rowMeans(thisdata[, sampleannotation[, "covariate"]=="M"]) 
meanRS = rowMeans(thisdata[, sampleannotation[, "covariate"]=="RS"]) 
meanDP = rowMeans(thisdata[, sampleannotation[, "covariate"]=="DP"])
meanN = rowMeans(thisdata[, sampleannotation[, "covariate"] =="N"]) 
rm(thisdata)
foldchange_RS_vs_M = meanRS - meanM
foldchange_RS_vs_M = foldchange_RS_vs_M[meanM>6 | meanRS>6]
top1000_RS_vs_M = names(foldchange_RS_vs_M[order(abs(foldchange_RS_vs_M), decreasing=TRUE)])[1:1000]
table(table_s2$ID %in% top1000_RS_vs_M) #  851 out of 909. For some reason it is not a 100% match.
```
Out of Table S2's **`r nrow(table_s2)`** probes, **`r sum(table_s2$ID %in% top1000_RS_vs_M)`** were reproduced. Scatter plots of the actual values shows a correlation.

@@

```{r dev='png', fig.width=8, fig.height=8, fig.show='hold', tidy=FALSE}
group = factor(sampleannotation$covariate)
block = as.factor(sampleannotation$chip)
design = model.matrix(~0+group+block)
fit = lmFit(datamatrices[["real_qnorm"]], design)
cont.matrix = makeContrasts ( contrasts="groupDP-groupN", levels=design)  
fit2 = contrasts.fit(fit, cont.matrix)
limma_ret_woc = eBayes(fit2)
limma_p_woc = limma_ret_woc$p.value
names(limma_p_woc) = names(limma_ret_woc$Amean)
rm(design, group, block, fit, cont.matrix, fit2)
table(p.adjust(limma_p_alt , method="fdr")<0.05, p.adjust(limma_p_woc , method="fdr")<0.05, dnn=c("ComBat adj", "LIMMA adj"))
```
The number of significant probes using the given FDR threshold of 0.05 when batch effect is handled by LIMMA is 
**`r sum(p.adjust(limma_p_woc , method="fdr")<0.05)`**. This is much lower compared to when ComBat is applied, 
**`r sum(p.adjust(limma_p_alt , method="fdr")<0.05)`**.  The p-value distribution is still skewed, but not as much.
```{r dev='png', fig.width=8, fig.height=8, fig.show='hold', tidy=FALSE}
hist(limma_p_alt, breaks=100, freq=T, border=thiscolors[1], main="Reproduced ANOVA p-values for DP vs. N")
hist(limma_p_woc, breaks=100, add=T, freq=T, border=thiscolors[2])
legend("topright", legend=c("ComBat adjusted", "LIMMA adjusted"), text.col=thiscolors)
```
